\systemEnding: Bootstrapping Domain-Specific Languages for Neurally-Guided Bayesian Program Learning

### 2018-02-09 paper.tex
small cleanups to notation and conclusion

We frame this problem as maximum a posteriori (MAP) inference in the
generative model diagrammed by Fig.~\ref{graphicalModel}(a). Writing $J$ for the joint probability of $(\mathcal{D},\theta)$, we want the $\mathcal{D}^*$ and $\theta^*$ solving:
\begin{align}\label{intractableObjectives}
\nonumber  J(\mathcal{D},\theta)&\triangleq \probability[\mathcal{D}]\probability(\theta)\prod_{x\in X} \sum_p \probability[x|p]\probability[p|\mathcal{D},\theta]\\
  \mathcal{D}^* &= \argmax_{\mathcal{D}}\int J(\mathcal{D},\theta)\;\mathrm{d}\theta \\
\nonumber  \theta^*& =\argmax_\theta J(\mathcal{D}^*,\theta)
\end{align}
The above equations summarize the problem from the point of view of an ideal Bayesian learner.
However, Eq.~\ref{intractableObjectives}
is wildly intractable because evaluating $J(\mathcal{D},\theta)$ involves
summing over the  infinite set of all possible programs.
In practice we will only ever be able to sum over a finite set of programs.
So, for each task, we define a finite set of programs, called a \emph{frontier}, and only marginalize over the frontiers:
\\\noindent\textbf{Definition.} A \emph{frontier of task $x$}, written $\mathcal{F}_x$,
is a finite set of programs where $\probability[x|p] > 0$ for all $p\in \mathcal{F}_x$.

Using the frontiers we  define the following intuitive lower bound on the joint probability, which we call $\lowerBound$:
\begin{align}
 J\geq \lowerBound\triangleq\probability[\mathcal{D}]\probability[\theta]\prod_{x\in X} \sum_{p\in \mathcal{F}_x} \probability[x|p]\probability[p|\mathcal{D},\theta]
\end{align}

@@

 \section{Contribution and Outlook}
 We contributed a model, \systemEnding, that learns to program by
 bootstrapping a DSL with new domain-specific primitives that the algorithm itself discovers.
  Alongside this DSL, we train a bottom-up
 neural recognition model that learns how to efficiently deploy the
 DSL and hone in on the right programs.
 %% We believe that this
 %% integration of top-down symbolic representation and bottom-up neural
 %% networks -- both of them learned -- is a useful general approach for
 %% getting program induction off the ground. 
Many directions remain open for future work. Two immediate goals are to
integrate more sophisticated neural recognition models~\cite{devlin2017robustfill} and
program synthesizers~\cite{solar2008program}, which may improve performance in some
domains over the simple, generic methods used here, and to explore DSL
meta-learning: Can we find a \emph{single} universal primitive set that could
effectively bootstrap DSLs for new domains, including the four domains
considered but many others?

>>>>

We frame this problem as maximum a posteriori (MAP) inference in the
generative model diagrammed by Fig.~\ref{graphicalModel}(a). Writing $J$ for the joint probability of $(\mathcal{D},\theta)$ and $X$, we want the $\mathcal{D}^*$ and $\theta^*$ solving:
\begin{align}\label{intractableObjectives}
\nonumber  J(\mathcal{D},\theta)&\triangleq \probability[\mathcal{D},\theta]\prod_{x\in X} \sum_p \probability[x|p]\probability[p|\mathcal{D},\theta]\\
  \mathcal{D}^* &= \argmax_{\mathcal{D}}\int J(\mathcal{D},\theta)\;\mathrm{d}\theta \\
\nonumber  \theta^*& =\argmax_\theta J(\mathcal{D}^*,\theta)
\end{align}
The above equations summarize the problem from the point of view of an ideal Bayesian learner.
However, Eq.~\ref{intractableObjectives}
is wildly intractable because evaluating $J(\mathcal{D},\theta)$ involves
summing over the  infinite set of all possible programs.
In practice we will only ever be able to sum over a finite set of programs.
So, for each task, we define a finite set of programs, called a \emph{frontier}, and only marginalize over the frontiers:
\\\noindent\textbf{Definition.} A \emph{frontier of task $x$}, written $\mathcal{F}_x$,
is a finite set of programs where $\probability[x|p] > 0$ for all $p\in \mathcal{F}_x$.

Using the frontiers we  define the following intuitive lower bound on the joint probability, which we call $\lowerBound$:
\begin{align}
 J\geq \lowerBound\triangleq\probability[\mathcal{D},\theta]\prod_{x\in X} \sum_{p\in \mathcal{F}_x} \probability[x|p]\probability[p|\mathcal{D},\theta]
\end{align}

@@

 \section{Contribution and Outlook}
 We contributed a model, \systemEnding, that learns to program by
 bootstrapping a DSL with new domain-specific primitives that the algorithm itself discovers.
  Alongside this DSL, we train a bottom-up
 neural recognition model that learns how to efficiently deploy the
 DSL and hone in on the right programs.
 %% We believe that this
 %% integration of top-down symbolic representation and bottom-up neural
 %% networks -- both of them learned -- is a useful general approach for
 %% getting program induction off the ground. 
Many directions remain open for future work. Two immediate goals are to
integrate more sophisticated neural recognition models~\cite{devlin2017robustfill} and
program synthesizers~\cite{solar2008program}, which may improve performance in some
domains over the simple, generic methods used here, and to explore DSL
meta-learning: Can we find a single universal primitive set that could
effectively bootstrap DSLs for new domains, including the four domains
considered but many others?
